---
title: "Filtering correlations"
output: html_notebook
---

Quick notes about looking at all preselected demographic variables.

```{r}
library(tidyverse)
library(here)
library(glue)
library(DBI)


```

```{r}
processed_db <- dbConnect(duckdb::duckdb(), here("4. Data","processed_data.duckdb"),read_only=FALSE)

```

## Source tables:

```{r}
dbListTables(processed_db)
```

```{r}
primary_vote <- 
tbl(processed_db,"primary_vote") |>
  select(Year,census_years,StateAb,DivisionNm,PartyAb,Percentage_Diff_National)
```

```{r}

```

## Age

```{r}
gen_order <- c("Age - Greatest Gen","Age - Silent Gen","Age - Baby Boomers","Age - Gen X","Age - Gen Y","Age - Gen Z","Age - Gen Alpha")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
tbl(processed_db,"age_and_sex") |>
  filter(Year!=2021) |>
  select(Year,Unit,Attribute,Percentage) |>
  collect() |>
  pivot_wider(names_from="Attribute",values_from = "Percentage") |>
  select(-Unit,-Year) |>
  relocate(any_of(gen_order)) |>
  GGally::ggpairs()
  
```

```{r}

tbl(processed_db,"age_and_sex") |>
  filter(Year!=2021) |>
  ggplot(aes(x=Percentage,y=fct_relevel(Attribute,gen_order))) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )
  #geom_histogram()+
  #facet_grid( Year ~ fct_relevel(Attribute,c("Age - Silent Gen")))

```

```{r}
tbl(processed_db,"primary_vote") |>
  select(-Percentage) |>
  left_join(tbl(processed_db,"age_and_sex"), by=c("DivisionNm"="Unit","census_years"="Year")) |>
  filter(census_years!=2021)|>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(fct_relevel(Attribute,gen_order)~PartyAb)+
  labs(x="Population Percentage",y="Primary Vote - % from National average")
  
  
```

```{r}

tbl(processed_db,"primary_vote") |>
  select(-Percentage) |>
  left_join(tbl(processed_db,"age_and_sex"), by=c("DivisionNm"="Unit","census_years"="Year")) |>
  filter(census_years!=2021)|>
  filter(str_detect(Attribute,"Gen Z")) |>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Year~PartyAb)+
  labs(x="Population Percentage",y="Primary Vote - % from National average",
       title= "Focus on Gen Z")


```

-   Greatest Gen, merge with silent Gen. - disappearing so problaby not so relevant in terms of changes.
-   focus on 4 Gens - from boomers to Zers.
-   Gen Alpha does not vote yet - Their influence can be captured through houselhold structure.
-   Very negative correlation
-   Note slighlty positive correlation between COAL vote and percentage of baby boomers, negative correlation with Labor.
-   Positive correlation between green and millenial vote, but not wih Zs . Be aware that Z percentage may not be relevant in earlier elections, as they were not of voting age (non linear model?)

## citizenship

Caution - Only citizens vote - this is just a proxy measure for multicultarity

```{r}
tbl(processed_db,"citizenship") |>
  filter(Year!=2021) |>
  ggplot(aes(x=Australian.Citizens,y="Citizens")) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )


```

```{r}
tbl(processed_db,"primary_vote") |>
  select(-Percentage) |>
  left_join(tbl(processed_db,"age_and_sex"), by=c("DivisionNm"="Unit","census_years"="Year")) |>
  filter(census_years!=2021)|>
  filter(str_detect(Attribute,"Gen Z")) |>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Year~PartyAb)+
  labs(y="Population Percentage",x="Primary Vote - % from National average",
       title= "Focus on Gen Z")
```

-   it looks like a big migration boom took place between 2011 and 2016.
-   Clear negative correlation between % of citizens and green vote. Not so clear with ALP/Coalition.
-   Not sure how informative this is, considering policy of encouraging acquisition of citizenship

## country of birth

Another proxy for multiculturality

```{r echo=FALSE, message=FALSE, warning=FALSE}
tbl(processed_db,"country_of_birth") |>
  filter(Year!=2021) |>
  select(-Unit,-Year) |>
  relocate(any_of(gen_order)) |>
  GGally::ggpairs()
  
```

```{r}
tbl(processed_db,"country_of_birth") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage") |>
  ggplot(aes(x=Percentage,y=fct_relevel(Attribute,gen_order))) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )
```

```{r}
tbl(processed_db,"primary_vote") |>
  select(-Percentage) |>
  collect() |>
  left_join(tbl(processed_db,"country_of_birth") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage"),
  by=c("DivisionNm"="Unit","census_years"="Year")) |>
  filter(census_years!=2021)|>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Attribute~PartyAb)+
  labs(y="Population Percentage",x="Primary Vote - % from National average")
```

-   Not sure how informative this is.
-   Not counting for 2nd people, which may still vote with cultural group.

## Language

```{r echo=FALSE, message=FALSE, warning=FALSE}
tbl(processed_db,"language") |>
  filter(Year!=2021) |>
  select(-Unit,-Year) |>
  GGally::ggpairs()
```

```{r}
tbl(processed_db,"language") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage") |>
  mutate(Attribute=str_remove_all(Attribute,"^Language...")) |>
  ggplot(aes(x=Percentage,y=Attribute)) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )


```

```{r}
primary_vote |>
  collect() |>
  left_join(tbl(processed_db,"language") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage"),
  by=c("DivisionNm"="Unit","census_years"="Year")) |>
  mutate(Attribute=str_remove_all(Attribute,"^Language...")) |>
  filter(census_years!=2021)|>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Attribute~PartyAb)+
  labs(y="Population Percentage",x="Primary Vote - % from National average")
```

-   Speakers of Aboriginal languagues are only significant in Lingiari (30%), then they are \~5% for Leichhardt, then MOSTLY \<1%

-   South Asian , Chinese on the rise, Arabic also relevant on some areas. More important is perhaps % of monolingual English speakers.

-   Chinese language seen positivelt correlated to COAL vote, South Asian Languages to ALP.

-   Cone behaivour for ALP with English monolingual percentages.

-   Highest correlations are negative - between English monolingual and Arabic, Chinese and East Asian languages

## Religion

```{r echo=FALSE, message=FALSE, warning=FALSE}
tbl(processed_db,"religion") |>
  mutate(across(contains("Religion"),as.numeric)) |>
  filter(Year!=2021) |>
  select(-Unit,-Year) |>
  GGally::ggpairs()
```

```{r}
tbl(processed_db,"religion") |>
  filter(Year!=2021) |>
  collect() |>
  mutate(across(contains("Religion"),as.numeric)) |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",
               values_to = "Percentage") |>
  mutate(Attribute=str_remove_all(Attribute,"^Religion...")) |>
  ggplot(aes(x=Percentage,y=Attribute)) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )

```

```{r}
primary_vote |>
  collect() |>
  left_join(tbl(processed_db,"religion") |>
  mutate(across(contains("Religion"),as.numeric)) |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage"),
  by=c("DivisionNm"="Unit","census_years"="Year")) |>
  mutate(Attribute=str_remove_all(Attribute,"^Religion...")) |>
  filter(census_years!=2021)|>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Attribute~PartyAb)+
  labs(y="Population Percentage",x="Primary Vote - % from National average")
```

-   Interesting to look at religion

-   anglican-prebysterian-uniting suburbs leaning towards Coalition

-   Catholic, Orthodox, Hindu, Islamic areas leaning towards ALP

-   Slight preference for greens by secular groups

-   No strong but positive correlation between non-Christian religions

## income

```{r echo=FALSE, message=FALSE, warning=FALSE}
tbl(processed_db,"income_level") |>
  filter(Year!=2021) |>
  select(-Unit,-Year) |>
  GGally::ggpairs()

```

```{r}

tbl(processed_db,"income_level") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",
               values_to = "Percentage") |>
  mutate(Attribute=str_remove_all(Attribute,"^Income...")) |>
  ggplot(aes(x=Percentage,y=Attribute)) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )
```

```{r}
primary_vote |>
  collect() |>
  left_join(tbl(processed_db,"income_level") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage"),
  by=c("DivisionNm"="Unit","census_years"="Year")) |>
  mutate(Attribute=str_remove_all(Attribute,"^Income...")) |>
  filter(census_years!=2021)|>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Attribute~PartyAb)+
  labs(y="Population Percentage",x="Primary Vote - % from National average")
```

-   income levels are highly correlated!

-   non clear associations with vote levels - except for negative association between greens and lower income range

## Household tenure

```{r echo=FALSE, message=FALSE, warning=FALSE}
tbl(processed_db,"household_tenure") |>
  filter(Year!=2021) |>
  select(-Unit,-Year) |>
  GGally::ggpairs()

```

```{r}
tbl(processed_db,"household_tenure") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",
               values_to = "Percentage") |>
  ggplot(aes(x=Percentage,y=Attribute)) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )
```

```{r}
primary_vote |>
  collect() |>
  left_join(tbl(processed_db,"household_tenure") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage"),
  by=c("DivisionNm"="Unit","census_years"="Year")) |>
  filter(census_years!=2021)|>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Attribute~PartyAb)+
  labs(y="Population Percentage",x="Primary Vote - % from National average")
```


-   more flat, less houses (obviously)

-   percentage of flats correlated to percantage of rentals
-  flat dwellers correlated to greens vote

## Education


```{r echo=FALSE, message=FALSE, warning=FALSE}
tbl(processed_db,"school_level") |>
  filter(Year!=2021) |>
  select(-Unit,-Year) |>
  GGally::ggpairs()

```
```{r}
tbl(processed_db,"school_level") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",
               values_to = "Percentage") |>
  ggplot(aes(x=Percentage,y=Attribute)) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )
```

```{r}
primary_vote |>
  collect() |>
  left_join(tbl(processed_db,"school_level") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage"),
  by=c("DivisionNm"="Unit","census_years"="Year")) |>
  filter(census_years!=2021)|>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Attribute~PartyAb)+
  labs(y="Population Percentage",x="Primary Vote - % from National average")
```




- merge Bachelor + graduate diploma + postgrad (ie. Uni education)
- explosion in certificates in the last decade - interisting
- perhaps make three categories - uni, tafe, other
- greens only evident correlation (pos Uni, negative tafe)

## Relationship

```{r echo=FALSE, message=FALSE, warning=FALSE}
tbl(processed_db,"relationship") |>
  filter(Year!=2021) |>
  select(-Unit,-Year) |>
  GGally::ggpairs()

```

```{r}
tbl(processed_db,"relationship") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",
               values_to = "Percentage") |>
  ggplot(aes(x=Percentage,y=Attribute)) +
  geom_violin() +
  facet_grid(.~Year) +
  theme(axis.title.y = element_blank()) +
   scale_x_continuous("Percentage of Electorate's Population"
                      )
```


```{r}
primary_vote |>
  collect() |>
  left_join(tbl(processed_db,"relationship") |>
  filter(Year!=2021) |>
  collect() |>
  pivot_longer(-c(Unit,Year),names_to = "Attribute",values_to = "Percentage"),
  by=c("DivisionNm"="Unit","census_years"="Year")) |>
  filter(census_years!=2021)|>
  ggplot(aes(x=Percentage,y=Percentage_Diff_National,colour=as_factor(Year)))+
  geom_point()+
  facet_grid(Attribute~PartyAb)+
  labs(y="Population Percentage",x="Primary Vote - % from National average")
```


- relevant for green vote : families with children, vs. group living, de factor, alone
- probably correlated with age

## Selection variables, single model

- focus only on top three


### disconnect DB

```{r}
dbDisconnect(processed_db,shutdown=TRUE)

```
