---
title: "Data Exploration"
format: html

---

```{r}
library(tidyverse)
library(auspol)
library(auscensus)
library(aussiemaps)
library(cropgrowdays)
library(rhandsontable)
library(here)
library(patchwork)


state_acronyms <- c("NSW","QLD","NT","WA","SA","VIC","TAS","ACT")
election_years <- as.character(c(2007,2010,2016,2022))
census_years <-c(2006,2011,2016,2021)

year_equivalency <- tibble(election_years=election_years,census_years=census_years)


base_theme <- customthemes::custom_plot_theme_md(google_font = "Roboto",background_colour = "white", 
                              legend_pos = "botom",legend_dir = "horizontal") +
            theme(legend.title = ggtext::element_markdown(size=6),
                  plot.title = ggtext::element_markdown(size=8),
                  legend.text = ggtext::element_markdown(size=6))

map_theme <- base_theme +
             theme(axis.title = element_blank(),
                   axis.ticks = element_line(colour="grey"))

#default theming for plots ####
theme_set(base_theme) 

customthemes::set_plot_colours(line_colour ="#0085c7",
                               fill_colour = tinter::lighten("#0085c7",0.5),
                               HVline_colour ="grey90")

```


This notebook looks at what sort of variables are available for to use.Variables documented here are used in the script that generates the analysis dataset. 

## Response Variables : Primary voting


Available election years


```{r}

auspol::list_years()

```

Get parties, coalesce coalition, greens

```{r}



parties <- auspol::list_parties() |>
  select(any_of(c(election_years,"PartyAb","PartyNm"))) |>
  pivot_longer(election_years,names_to = "Year",values_to="Value") |>
  filter(Value) |>
  distinct(PartyAb,PartyNm) 
  
  
parties |>
   rhandsontable(readOnly = TRUE) |>
    hot_cols(columnSorting = TRUE)

write_csv(parties,here("3. Data Extraction","parties.csv"))

```


```{r}
parties <- read_csv(here("3. Data Extraction","parties_edited.csv"))

parties <-parties |> select(Grouping,PartyAb) |> 
  group_by(Grouping) |> summarize(Parties=list(unique(PartyAb)))

parties_merge <-list()

for(i in 1:nrow(parties)){
  parties_merge[length(parties_merge)+1] <- parties[i,]$Parties 
}
names(parties_merge) <- parties$Grouping

primary_vote <- auspol::house_primary_vote_summary(state= state_acronyms, 
                                   parties=names(parties_merge),
                                   merge_parties = parties_merge,
                                   include_names=FALSE,
                                   year=election_years) |>
  select(Year,StateAb,DivisionNm,PartyAb,OrdinaryVotes,Percentage)


primary_vote |> 
  rhandsontable(readOnly = TRUE) |>
    hot_cols(columnSorting = TRUE)
```

```{r}

party_palette <- auspol::manage_colours(extra_values = unique(primary_vote$PartyAb),
                                        palette =ochRe::ochre_palettes[["tasmania"]])

primary_vote |>
  ggplot(aes(x=PartyAb,y=Percentage,colour=PartyAb,fill=PartyAb)) +
  ggbeeswarm::geom_beeswarm(alpha=0.5) +
  facet_grid(Year ~ StateAb) +
  scale_color_manual(values=party_palette)+
  scale_fill_manual(values=party_palette) +
  theme(axis.text.x = element_text(angle=90,hjust=1)) 

```


## Explanatory variables


### Census Data

Type of data - census questions

```{r}

census_tables <- auscensus::list_census_tables() |>
  as.data.frame()


census_tables|>
   flextable::flextable()

```
Interesting tables:

* 04	Age by Sex
* 09	Country of Birth of Person by Age by Sex
* 08	Ancestry by Birthplace of Parents

* 39 Non-School Qualification: Level of Education by Age by Sex
* 40 Non-School Qualification:  Level of Education by Age by Sex
* 46 Non-School Qualification:  Level of Education by Age by Sex
* 49 Highest Non-School Qualification:  Level of Education by Ag


*13 Language Used at Home by Proficiency in Spoken English by Sex
*12 Language Spoken at Home by Sex
*13 Language Spoken at Home by Sex
*13 Language Spoken at Home by Proficiency in Spoken English

* 16	Gross Individual Income (Weekly) by Age by Sex	- changed from previous census
* 17	Total Personal Income (Weekly) by Age by Sex




* 18 Voluntary Work for an Organisation or Group by Age by Sex
* 19 Voluntary Work for an Organisation or Group by Age by Sex
* 23 Voluntary Work for an Organisation or Group by Age by Sex


Extract values per electorate and then ask about differences within electorates


### 04 Age by Sex

```{r}

# Form list get groups of variables, chunked by generation
var04 <- list_census_attributes(number="04")

var04 <- var04 |>
  filter(str_detect(Attribute,"\\d{1,3}_\\d{1,3}_",TRUE)) |>
  filter(str_detect(Attribute,"\\d{1,3}-\\d{1,3}",TRUE))  |>
  filter(str_detect(Attribute,"\\d{1,3}_persons$",TRUE))  |>
   filter(str_detect(Attribute,"[Oo]ver",TRUE))  |>
   filter(str_detect(Attribute,"[Tt]otal.*[Mm]ales",TRUE)) |>
  pivot_longer(c(-Table,-Attribute), names_to="Year",values_to = "Value") |>
  mutate(Year=as.numeric(Year)) |>
  filter(!is.na(Value)) |>
  mutate(Age_at_census=str_extract(Attribute,"\\d{1,2}") |> as.numeric(),
         YOB = Year-Age_at_census,
         Gender=case_when(
           str_detect(Attribute, "_[Mm]ales$") ~ "Males",
           str_detect(Attribute, "_[Ff]emales$") ~ "Females",
           TRUE ~"-"))

# 80 means 80 and over!
generations <- bind_rows(tibble(YOB=1883:1900,Group="Lost Gen"),
                         tibble(YOB=1901:1927,Group="Greatest Gen"),
                         tibble(YOB=1928:1945,Group="Silent Gen"),
                         tibble(YOB=1946:1964,Group="Baby Boomers"),
                         tibble(YOB=1965:1980,Group="Gen X"),
                         tibble(YOB=1981:1996,Group="Gen Y"),
                         tibble(YOB=1997:2012,Group="Gen Z"),
                         tibble(YOB=2013:2030,Group="Gen Alpha"))

var04 <- var04 |>
        left_join(generations,by="YOB")

generations_list <- unique(generations$Group)
attributes_list <- list()
i<-1

for(year in census_years){
  attributes_i <- list()
  j <- 1

for(gen in generations_list){

  for(gender in c("Females","Males")){
    
  grouping <- var04 |> 
    filter(Group==gen & Year==year & Gender==gender) |>
    distinct(Attribute) |>
    pull()

  naming <- str_c(gen," - ",gender)
  
  if(length(grouping>0)){
     attributes_i[[j]] <- grouping
     names(attributes_i)[j] <- naming
     j <- j + 1
    
  }  
  }
}
  
  attributes_list[[i]] <- attributes_i
  names(attributes_list)[i] <- year
  i <- 1+i
  
}  
rm(grouping,naming,attributes_i,i,j,gen,year,generations_list)

#totals for percentage calculations

totals <- var04   |> 
          filter(is.na(Age_at_census)) |>
          distinct(Attribute) |>
          pull()
totals <- list("Total"=as.vector(totals))


# get the census data

age_and_sex <- tibble()

for(x in census_years){

  if(x=="2006"){
    geo_structure_x <- "CED_2007"
  } else{
    geo_structure_x <- "CED"
  }
  
censusx <- get_census_summary(table_number = "04",
                      geo_structure = geo_structure_x,
                      selected_years = as.character(x),
                      attribute = attributes_list[[which(names(attributes_list)==x)]],
                      reference_total = totals) |>
        filter(str_detect(Unit,"Shipping",TRUE))|>
        filter(str_detect(Unit,"Usual",TRUE))   |>
        filter(str_detect(Unit,"Not Applicable",TRUE)) 
  


age_and_sex <- bind_rows(age_and_sex,censusx)

}

rm(censusx,x,var04,totals)



```


```{r}
plots <- list()
for(i in 1:length(census_years)){
  
  plots[[i]] <- age_and_sex |>
                filter(Year==census_years[i])           |>
                mutate(Electorate=str_c(Unit,"-",Year)) |>
                ggplot(aes(x=as.character(Year),y=Percentage,label=Electorate)) +
                ggbeeswarm::geom_beeswarm() + 
                facet_wrap(Attribute ~., ncol = 2 ) 
}

patchwork::wrap_plots(plots,nrow=1)


```

```{r}

primary_vote <- primary_vote |>
                select(-OrdinaryVotes)|>
                pivot_wider(names_from = "PartyAb",values_from = "Percentage") |>
                mutate(across(where(is.numeric), ~ if_else(is.na(.x),0,.x)))   

age_and_sex <- age_and_sex |>
                select(Unit,Year,Attribute,Percentage) |>
                pivot_wider(names_from = "Attribute",values_from = Percentage)|>  
                left_join(year_equivalency,by=c("Year"="census_years")) |>
                mutate(Year=as.numeric(election_years))

 

```

```{r}
primary_vote |>
  left_join(age_and_sex, by=c("DivisionNm"="Unit","Year"="Year")) |>
  select(-election_years) |>
  mutate(across(where(is.numeric), ~ if_else(is.na(.x),0,.x)))  |> 
  pivot_longer(-c(Year,StateAb,DivisionNm,ALP,COAL,GRN,`Other Left`,`Other Right`,Other),
               names_to="Population Group",
               values_to="Pop Percentage")|>
  pivot_longer(c(ALP,COAL,GRN,Other,`Other Left`,`Other Right`),
               names_to = "Party",
               values_to = "Party Percentage") |>
  ggplot(aes(x=`Pop Percentage`,y=`Party Percentage`)) +
  geom_point(alpha=0.5) +
  facet_grid(Party ~`Population Group`)
```




### Non School Qualification 

```{r}
census_tables <- list_census_tables() |>
                 filter(Number %in% c("39","40","46","49")) |>
                 filter(str_detect(`Table Name`,"Non-School")) |>             
                 filter(str_detect(`Table Name`,"Level of Education by Age by Sex"))


tables_summary <- census_tables |>
                  select(-contains("Table")) |>
                  pivot_longer(-Number, names_to="Year",values_to = "Value") |>
                  filter(!is.na(Value)) |>
                  select(-Value)

school_attributes <- tibble()

for(i in 1:nrow(tables_summary)){
  
  table_i <- tables_summary[i,]$Number
  year_i  <- tables_summary[i,]$Year
  
  attr_i <- list_census_attributes(table_i) |> 
             pivot_longer(-c(Table,Attribute),
                           names_to="Year",values_to = "Value") |>
              filter(Year==year_i) |>
              filter(!is.na(Value)) |>
              select(-Value)
  
  school_attributes <- bind_rows(attr_i,school_attributes)
  
}

  school_attributes <- school_attributes   |> 
                       distinct(Attribute) |>
                      filter(str_detect(Attribute,"[Pp]erson")) |>
                      filter(str_detect(Attribute,"[Aa]ge",TRUE)) |>
                      filter(str_detect(Attribute,"[Yy]ear",TRUE))
school_attributes |> 
  rhandsontable(readOnly = TRUE) |>
  hot_cols(columnSorting = TRUE)

write_csv(school_attributes,"ed_levels.csv")
  
```

```{r}
school_attributes <- read_csv("ed_levels_edited.csv")

totals <- list(Total = school_attributes |> filter(Attr_New=="Total") |> pull(Attribute))

school_levels <- school_attributes |> filter(Attr_New!="Total") |> distinct(Attr_New) |> pull()

levels <- list()
i <-1
for(level in school_levels){
  
  levels[[i]]      <-  school_attributes |> filter(Attr_New==level)  |> pull(Attribute)
  names(levels)[i] <-  level
  i <- i +1
  
  
} 

school_level <- tibble()

for(year in census_years){
  
  if(year=="2006"){
    geo_structure_x <- "CED_2007"
  } else{
    geo_structure_x <- "CED"
  }
  
  table_number <- tables_summary |> filter(Year==year) |> pull(Number)
  
  school_level_i <- get_census_summary(census_table = census_tables,
                                       selected_years = year,
                                       geo_structure=geo_structure_x,
                                       attribute = levels,
                                       reference_total = totals) |>
  filter(str_detect(Unit,"Shipping",TRUE))|>
  filter(str_detect(Unit,"Usual",TRUE))   |>
  filter(str_detect(Unit,"Not Applicable",TRUE)) 
  
  school_level <- bind_rows(school_level,school_level_i)
  
}
b<-get_census_data(census_tables,"CED")



```
Table does not exist for CED in 2011 - we need to aggreate from SA1s

## Climate Data
