--- 
title           : "Analysing Australian Electoral Results"
author          : "Carlos Yanez Santibanez"
programme       : "Data Analytics MSc (ODL)"
subject         : "Project"
date            : 26 June 2023
output: pdf_document
params:
  book: FALSE
  html: TRUE
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(readr)
library(stringr)
library(here)
library(auspol)
library(ggplot2)
library(tidyr)
library(fs)
library(tidymodels)
library(vip)
library(glue)
library(kableExtra)
library(flowdiagramr)
#flextable::add_latex_dep(TRUE,TRUE)
source(here::here("formatting_defaults.R"))

project_path <- fs::path(here(),"..")

knitr::opts_knit$set(message=FALSE,echo=FALSE,warning=FALSE,cache=FALSE)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)

eval_book <- params$book
eval_html <- params$html

if(eval_html){
  library(leaflet)
  library(htmltools)
  library(plotly)
  library(crosstalk)
}

keep_vars <- c(ls(),"keep_vars")


```

```{r eval=eval_book}
uofgdataanalyticsreport::cd_page_title()
```
 

```{r eval=FALSE, include=FALSE}
bookdown::render_book("index.Rmd", "bookdown::gitbook",params=list(book=FALSE,html=TRUE))
bookdown::render_book("index.Rmd", "bookdown::pdf_book",params=list(book=TRUE,html=FALSE))


```


`r if(eval_book)'\\cleardoublepage \\pagenumbering{roman}'`




`r if(eval_html) '# Dedication {-}'`

`r if(eval_book) '\\clearpage \\begin{raggedright} \\thispagestyle{empty} \\vspace*{\\fill}'` `r if(eval_html) '<br><br><center><b><i>'`
To Lauren. Thanks for you constant, never-ending support, encouragement and patience.
 
To you and Lucas, with whom I have made Australia home. 
`r if(eval_html) '</center></b></i>'` `r if(eval_book) '\\vspace*{\\fill} \\end{raggedright} \\clearpage'`

\newpage

```{r eval=eval_book}
uofgdataanalyticsreport::cd_toc()
```
