---
title           : "Data Summary"
programme       : "Data Analytics MSc (ODL)"
subject         : "Project"
author          : "Carlos Yanez Santibanez"
date            : "15 October 2022"

documentclass   : article   # article (default) report
fontsize        : 11pt      # 10pt 11pt (default) 12pt
lang            : en-GB
output:
  bookdown::pdf_book:
    base_format: uofgdataanalyticsreport::uofgdataanalyticsreport
    pandoc_args: --listings
    fig_caption: yes
    number_sections: yes
    toc: no
    md_extensions: +yaml_metadata_block
---

```{r setup, include = FALSE}

library(tidyverse)
library(here)
library(fs)
library(uofgdataanalyticsreport)
library(lubridate)
library(showtext)
library(sysfonts)
library(sf)
library(knitr)
library(kableExtra)

sysfonts::font_add_google("Roboto","Roboto")
showtext::showtext_auto()

knitr::opts_chunk$set(
 # fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  eval = TRUE,
  include=TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)


knitr::knit_hooks$set(inline = function(x) { if(!is.numeric(x)){ x }else{ digits_formatter(x,sig_digits = 3) } })

#vic_hex <- st_read(path(here(),"2. Meetings","1. Meeting","vic_hex.geojson"))
vic_hex <- st_read(path(here(),"2. Meetings","1. Meeting","vic-july-2021-esri","E_VIC21_region.shp")) %>%
                   st_zm()

```

```{r title_page, echo=FALSE,include=TRUE}
cd_header_footer()  #adds header and footer
cd_page_title()     #title page

#to customise, run ?function to see the relevant parameters
```

```{r load_data}

files_dir <- path(here(),"2. Meetings","1. Meeting","0. data sample")
files <-dir_ls(files_dir,regexp = "[.]csv$")

data <- list()
for(i in 1:length(files)){
  data[[i]] <- read_csv(files[i])
}

files <- str_remove_all(files,files_dir) %>% 
         str_remove_all(.,"[/]")         %>%
         str_remove_all(., "[.]csv")
  
names(data) <- files
rm(files_dir,files)

```

# Data Summary

The data here presented contains:

-   Electoral results for the Australian 2022 Federal Election.
-   Census 2021 Results
-   Other Statistical Information, retrieved by the {readabs} package.
-   Sample meteorological data, collected by the {cropgrowdays} package.

# Electoral Results

Electoral results are available on the Australian Electoral Commission's (AEC) Tally Room website (<https://results.aec.gov.au/27966/Website/HouseDefault-27966.htm>).

Data available shows the primary vote for every candidate in each polling station. As an example, table \@ref(tab:tab1) presents the results for two polling stations in the federal electoral division (electorates, divisions) of Macnamara (in Melbourne). Table \@ref(tab:tab2) shows an aggregate per party for the same electorate.

```{r tab1}
data$sample_fed_elections_vic_2022 %>% 
  filter(DivisionNm=="Macnamara") %>%
  filter(PollingPlace %in% c("Albert Park","Alfred Square")) %>%
  kable(caption="Milestones",booktabs = T) %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"),
                full_width = FALSE,
                position = "center") %>%
  row_spec(0,bold=TRUE)

```

```{r tab2}

votes_per_electorate <- data$sample_fed_elections_vic_2022 %>%
                        group_by(DivisionNm) %>%
                        summarise(total_votes = sum(OrdinaryVotes,na.rm=TRUE),.groups="drop")

primary_vote <- data$sample_fed_elections_vic_2022 %>%
                group_by(DivisionNm,PartyNm) %>%
                summarise(primary_vote = sum(OrdinaryVotes,na.rm=TRUE),.groups="drop") %>%
                left_join(votes_per_electorate,by="DivisionNm") %>%
                mutate(primary_perc = primary_vote/total_votes) %>%
                select(-total_votes)
  

primary_vote %>% 
    filter(DivisionNm=="Macnamara") %>%
    arrange(desc(primary_perc))     %>%
    kable(caption="Milestones",booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position","striped"),
                full_width = FALSE,
                position = "center") %>%
    row_spec(0,bold=TRUE)


```

This data is available for all electorates. Figure \@ref(fig:fig1) show primary voting for the Australian Labor Party (ALP) in Victoria.

```{r fig1}


primary_vote %>% 
    rename("Percentage"="primary_perc") %>%
    filter(PartyNm == "Australian Labor Party") %>%
    left_join(vic_hex,by=c("DivisionNm"="Elect_div")) %>%
    ggplot(aes(geometry=geometry,fill=Percentage,label=DivisionNm)) +
    geom_sf(colour=NA) +
    customthemes::custom_map_theme() +
    scale_fill_gradient(low="#FFFFFF",high="#E13940") +
    labs(title="Vote in Victoria - Labor Party")
  
  

```

# Census 2021 Results

The Australian Bureau of Statistics (ABS) publishes data packs with the Census data (<https://www.abs.gov.au/census/find-census-data/datapacks>). The ABS provides these data packs in a range of geographical aggregations, including federal electorates.

Data for this presentation has been retrieved using the {Census2021.DataPack} package (<https://github.com/HughParsonage/Census2021.DataPack>), but there is scope to improve the retrieval method.

Table \@ref(tab:tab3) shows a summary of all the data collated by the census. Table \@ref(tab:tab4) and figure \@ref(fig:fig2) show a sample of median income per household in Victorian electorates.

```{r tab3}
data$census_tables  %>%
  kable(caption="Census tables",booktabs = T) %>%
  kable_styling(latex_options = c("HOLD_position","scale_down", "striped"),
                full_width = FALSE,
                position = "center") %>%
  row_spec(0,bold=TRUE)

```

```{r tab4}
data$household_income %>%
  head() %>%
  kable(caption="Median House Income - Sample",booktabs = T) %>%
  kable_styling(latex_options = c("HOLD_position","stried"),
                full_width = FALSE,
                position = "center") %>%
  row_spec(0,bold=TRUE)

```

```{r fig2}
data$household_income %>% 
  rename("Median Income"="medianTotalHouseholdIncome") %>%
  right_join(vic_hex,by=c("CED_NAME21"="Elect_div")) %>%
   ggplot(aes(geometry=geometry,fill=`Median Income`)) +
    geom_sf(colour=NA) +
    customthemes::custom_map_theme() +
    scale_fill_gradient(low="purple",high="green") +
    labs(title="Median House Income in Victoria- 2021 Census")
  
```

# Other statistical data

The ABS also provide a range of other statistical data, through their Time Series and Data Cubes products. This data is not aggregated by electorate - it could be presented by Local Government Area (LGA), State or other.

For this type of data, the package {readabs} provide a good interface. This data can be used to supplement Census data if necessary.

# Weather Data

The Bureau of Meteorology (BoM) provides historical data since the start of the 20th century. This data is not directly available for programmatic download from the BoM's website - however the package {cropgrowdays} (<https://gitlab.com/petebaker/cropgrowdays/>) provides a way to retrieve national data from Queensland's government Longpaddock service (<https://www.longpaddock.qld.gov.au/>).

A sample for Melbourne is presented in table \@ref(tab:tab5).

```{r tab5}
data$sample_bom_mel %>%
  head() %>%
  kable(caption="Milestones",booktabs = T) %>%
  kable_styling(latex_options = c("HOLD_position"),
                full_width = FALSE,
                position = "center") %>%
  row_spec(0,bold=TRUE)

```
